---
title: "births_analysis"
author: "Anna Khoo"
date: "29 June 2020"
output: html_document
---

This is a script to look at the sector breakdown of new enterprises registered on Companies House by incorporation date.
It's a variant on the sic_match in Insolvency_master.Rmd, but runs off the latest snapshot on Companies House.
Because snapshots are very large (4m rows), this script takes a while to run and needs a clear environment to give enough RAM.

```{r setup, include=FALSE}
.libPaths("D:/R/Libraries")
pacman::p_load(tidyverse, janitor, data.table, zoo, ggplot2)
```


```{r load latest snapshot}
snapshot_sic <- fread("BasicCompanyDataAsOneFile-2020-07-01/BasicCompanyDataAsOneFile-2020-07-01.csv",
                  select=c(1:13,15,20,27:30,33))
```
We'll split by time now, because otherwise this stage has to be written in multiple times later on, it's very useful in all child objects.
```{r split date - moved further up before join}
snapshot_sic <- snapshot_sic %>%
  tidyr::separate(col=IncorporationDate, into=c("day","month","year"), sep="/")
```


Now we run the sic_match function, written in full here for readability. We'll write out the resulting object to save time on analysis later if we want to skip creating it every time.


```{r reshape to allow sic join}
snapshot_sic_only <- snapshot_sic %>%
  pivot_longer(cols=18:21, 
               names_to = "sicfield",
               values_to = "siccode",
               names_prefix= "SICCode.",
               values_drop_na=TRUE 
  )

#that's really big, so filter for siccodes that exist:
snapshot_sic_only <- snapshot_sic_only %>% 
  filter(!siccode=="")

snapshot_sic_only <- snapshot_sic_only %>%
  tidyr::separate(col=siccode, into=c("siccode","sicdesc"), sep=" - ")

snapshot_sic_only$sicdiv <- substr(snapshot_sic_only$siccode,1,2)
#import ref
sicdivs <- rio::import("sic2007_division_group_class.xls", sheet = 2)
colnames(sicdivs) <- c("sicdiv","divisionname")

snapshot_sic_only <- left_join(snapshot_sic_only, sicdivs)

industry_to_div <- industry_to_div <- readxl::read_excel("industry_to_div.xlsx", 
    sheet = "industry_to_div") %>% 
  select(-division_top)

industry_to_div$division_bottom <-  str_pad(industry_to_div$division_bottom,2,pad="0")
industry_to_div <- industry_to_div %>% 
  rename(sicdiv=division_bottom)

snapshot_sic_only <- left_join(snapshot_sic_only, industry_to_div)

###

write_csv(snapshot_sic_only, "snapshot_sic_only.csv")
```

#Run from csv possible from here:
```{r read in snapshot matched to sic data}
snapshot_sic_only <- read_csv("snapshot_sic_only.csv")
```
```{r deduplicate industry, plot by month/year}
#deduplicate industry desc over company number to avoid double counting where multiple sic codes match to the same broad sector:
snapshot_sic_new_industry  <- snapshot_sic_only [!duplicated(snapshot_sic_only[c(2,24)]),]
```

```{r pivot over time}
pivot_mth_yr <- snapshot_sic_new_industry %>%
  filter(year=="2019"|year=="2020") %>% 
  group_by(year, month, industry_desc) %>% 
  summarise(births=n()) %>% 
  arrange(month)

grouped_bars <- ggplot(pivot_mth_yr, 
                       aes(x = month, 
                           y = births, 
                           fill = year)) +
  facet_wrap(~industry_desc, scales="free")+
  geom_bar(stat="identity", position = position_dodge2(preserve = "single")) +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  scale_fill_manual(values = c("#1380A1", "#FAAB18","#8B0000"))+
  theme(axis.text.x = element_text(angle = 30, hjust = 1))+
  labs(title = "Enterprise births 2020", x="Month", y="Incorporated on CH")
grouped_bars
```
```{r write out pivot - raw is too big to load outside of R}
#using raw sheet NOT industry matched sheet for basic figures over time to avoid duplication over industry type per company
pivot_mth_yr <- snapshot_sic %>%
  filter(year=="2019"|year=="2020") %>% 
  group_by(year, month) %>% 
  summarise(births=n()) %>% 
  arrange(month)

write_csv(pivot_mth_yr, "births_over_time_core.csv")
```

```{r create normalised industry pivot for snapshot_sic_new}

lockdown = c("04", "05", "06")

lockdown_new <- snapshot_sic_new_industry %>% 
  filter(year==2020, month %in% lockdown)

# count new enterprises by industry category, as a proportion of all companies set up during lockdown months
# NROW counts nrow for a vector, here unique(x)

new_biz_by_industry <- lockdown_new %>% 
  count(industry_desc, sort=T) %>% 
  mutate(proportion=(n/NROW(unique(lockdown_new$CompanyNumber)))*100) %>% 
  adorn_totals("row")

write_csv(new_biz_by_industry, "new_biz_by_industry.csv")

#the total checks that the sum is >100 per cent, so including multiple industry counts
```
```{r plot new industry bar}
#now plot

new_biz_by_industry <- new_biz_by_industry %>% 
  filter(!industry_desc=="Total")

bars <- ggplot(new_biz_by_industry, 
                       aes(x=reorder(industry_desc,proportion), 
                           y=proportion)) +
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 30, hjust = 1))+
  labs(title = "Enterprise births 2020", x="Broad industry group", y="Proportion of new businesses registered since lockdown")+
  coord_flip()
bars

```
Interesting...what about sic division then, so one level of industry down:

```{r DON'T RUN moving this higher up in main script}
snapshot_sic_only <- snapshot_sic_only %>%
  tidyr::separate(col=IncorporationDate, into=c("day","month","year"), sep="/")
```
```{r fastest growing sic division?}

lockdown = c("04", "05", "06")

lockdown_new_div <- snapshot_sic_only %>% 
  filter(year==2020, month %in% lockdown)

#sub out sicdesc here for more granular look, retail shopping is highest, NSS

new_biz_by_div <- lockdown_new_div %>% 
  count(sicdesc, sicdiv, sort=T) %>% 
  mutate(proportion=(n/NROW(unique(lockdown_new_div$CompanyNumber)))*100) %>% 
  head(20) %>% 
  adorn_totals("row")

#now plot

new_biz_by_div <- new_biz_by_div %>% 
  filter(!sicdesc=="Total")

bars <- ggplot(new_biz_by_div, 
                       aes(x=reorder(sicdesc,proportion), 
                           y=proportion)) +
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 30, hjust = 1))+
  scale_colour_manual("#1380A1")+
  labs(title = "Enterprise births 2020", x="Industry division", y="Proportion of new businesses registered since lockdown")+
  coord_flip()
bars

write_csv(new_biz_by_div, "new_biz_by_desc.csv")
```

What about births by registered office address?
Using main snapshot_sic object here to avoid duplication issues in snapshot_sic_new_industry over industry type.
```{r attempt breakdown by area}
postcode_to_ons_ref <- fread("NSPL_FEB_2020_UK.csv", select=c("pcds","laua"))
postcode_to_ons_ref <- postcode_to_ons_ref %>% 
  rename(postcode=pcds)
snapshot_sic<- snapshot_sic %>%
  rename(postcode=RegAddress.PostCode)
snapshot_sic <- left_join(snapshot_sic, postcode_to_ons_ref)
```
```{r bind to la and pivot}
la_name_to_ons_code <- read_csv("Local_Authority_Districts_December_2019_Names_and_Codes_in_the_United_Kingdom.csv")

#names(la_name_to_ons_code)
la_name_to_ons_code <- la_name_to_ons_code %>%
  select(LAD19CD, LAD19NM) %>% 
  rename("laua"=LAD19CD, "LA_name"=LAD19NM)

#subset by time
lockdown = c("04", "05", "06")
lockdown_new <- snapshot_sic %>% 
  filter(year==2020, month %in% lockdown)

#join
lockdown_new <- left_join(lockdown_new, la_name_to_ons_code)

#check for non-matches
find_na_LA <- lockdown_new %>% 
  filter(is.na(LA_name))
#another one lost, others are carried nas from postcode match fail

#Pivot by area (but this is not particularly helpful, we need to normalise)
la_pivot <- lockdown_new %>% count(LA_name, sort=T)

#now 
#note this will overwrite version created by original script 'postcode_matching.R'
write_csv(la_pivot, "new_by_district.csv")
```

#Fixing NEC
n.e.c is distorting the entertainment bar, although technically included.
We'll find out how many there are of those and how many matches we might have to redo, sic range etc:
```{r find n.e.c}
find_nec <- lockdown_new_div %>%
  filter(str_detect(sicdesc, "n.e.c"))

find_90s <- lockdown_new_div %>% 
  filter(industry_desc=="Arts, entertainment, recreation & other services")
```
Found quite a few. Have now changed the industry_div match table to split Arts/E/R and 'other services' in lookup.
But we'll leave this here for record.


```{r closer look at June newbies, flatten}
June_new <- snapshot_sic_only %>% 
  filter(year=="2020", month=="06") %>%
  select(-siccode, -sicdiv, -sicdesc, -divisionname)

#deduplicate industry desc over company number
June_new <- June_new[!duplicated(June_new[c(2,20)]),]

industry_June_new <- June_new %>% 
  count(industry_desc, sort=T) %>% 
  mutate(proportion=n/nrow(unique(June_new$CompanyNumber)))

#to flatten, add this line:
#pivot_wider(names_from = sicfield,
              #values_from = industry_desc) %>%
#unite(industry_desc_all,19:22, sep = ",", remove = TRUE, na.rm = T)

  #mutate(industry_all = unique(sicfield))
  
```
#Quality checks - for record

The melting process may (possibly) have lost observations due to NA or blank in SicText.1 field in the pivot. We'll have a closer look:
```{r check no SicText.1 obs lost in value_drop_na and filter}
#this is high memory use, so careful now, check separately:

snapshot_sic_all <- snapshot_sic %>%
  pivot_longer(cols=16:19, 
               names_to = "sicfield",
               values_to = "siccode",
               names_prefix= "SICCode."
  )

snapshot_sic_blank <- snapshot_sic_all %>% 
  filter(sicfield=="SicText_1") %>% 
  filter(is.na(siccode))

snapshot_sic_blank <- snapshot_sic_all %>% 
  filter(sicfield=="SicText_1") %>% 
  filter(siccode=="")

snapshot_sic_blank <- snapshot_sic_all %>% 
  filter(sicfield=="SicText_1")

#same number of obs as non-pivoted parent, no obs being lost
```


We need to check what the drop-off rate is for companies, so if we've lost any companies in the most recent snapshot that were on the database as newly incorporated in January/Feb, for example:
```{r check backdated tally, load older snapshot}
May_sic <- fread("BasicCompanyDataAsOneFile-2020-05-01/BasicCompanyDataAsOneFile-2020-05-01.csv",
                  select=c(1:13,15,20,27:30,33))
```
```{r find older births}
May_sic_new <- May_sic  %>%
  tidyr::separate(col=IncorporationDate, into=c("day","month","year"), sep="/") %>% 
  filter(year==2020, (month=="03"| month=="02"|month=="01"))

#pivot and count:
pivot_mth_yr_may <- May_sic_new %>%
  group_by(year, month) %>% 
  summarise(births=n()) %>% 
  arrange(month)

pivot_mth_yr_new <- snapshot_sic %>%
  filter(year=="2020") %>% 
  group_by(year, month) %>% 
  summarise(births=n()) %>% 
  arrange(month)

check_losses <- left_join(pivot_mth_yr_new, pivot_mth_yr_may, by=c("year", "month"))
#by subbing in we can check further back:
#check_losses <- left_join(check_losses, pivot_mth_yr_apr, by=c("year", "month"))
#check_losses <- left_join(check_losses, pivot_mth_yr_mar, by=c("year", "month"))

```
Birth count by month is very stable looking back, not identical but if anything, there are a few more companies known to be incorporated in a given month in the newer dataset rather than in the older snapshot, probably due to backdating.
This means our births chart, kept to recent months, is a fair representation when run from the newest snapshot.

#Attempting by company size


Copying from time_analysis and adapting to find births by company size.
We're particularly interested in retail/wholesale/motor trading as those are the resurgence spikes.
```{r snapshot pivot over time by acc.category}

#we just want company size indication, clean col/subset:
snapshot_sic_new$Accounts.AccountCategory <- gsub("TOTAL EXEMPTION ", "",snapshot_sic_new$Accounts.AccountCategory)

accounts <- c("FULL","GROUP","MEDIUM","NO ACCOUNTS FILED","SMALL", "DORMANT")

pivot_mth_yr <- snapshot_sic_new %>%
  filter(year==2020) %>%
  filter(Accounts.AccountCategory %in% accounts) %>% 
  group_by(year, month, Accounts.AccountCategory, industry_desc) %>% 
  summarise(births=n()) %>% 
  arrange(month)


#pivot_mth_yr$month <- factor(pivot_mth_yr$month, levels= month.name)
pivot_mth_yr$year <- as.character(pivot_mth_yr$year)

grouped_bars <- ggplot(pivot_mth_yr, 
                       aes(x = month, 
                           y = births, 
                           fill = year)) +
  facet_wrap(~Accounts.AccountCategory, scales="free")+
  geom_bar(stat="identity", position = position_dodge2(preserve = "single")) +
  geom_hline(yintercept = 0, size = 1, colour="#333333") +
  scale_fill_manual(values = c("#1380A1", "#FAAB18","#8B0000"))+
  theme(axis.text.x = element_text(angle = 30, hjust = 1))+
  labs(title = "Enterprise births 2020", x="Month", y="Incorporated on CH")
grouped_bars
```
So we can't look at size of new enterprise according to accounts - they won't file yet. Accounting timetables have also been relaxed under the Insolvency Act 2020.

