---
title: "Insolvencies - dealing with detail pages"
output: html_notebook
---

# Insolvencies

In the first notebook we combined, cleaned and analysed scraped [corporate insolvency notices](https://m.thegazette.co.uk/insolvency?categorycode=G205010000&results-page=1) on the Gazette. 

We realised that scraping the detail pages for each notice was going to be needed. 

So, in this notebook, we will combine those scrapes and perform some cleaning and analysis.

NOTE: for convenience the files are in the 'gazettescrapeApr9' folder and need to be moved out before running the code below.

## Combining the data

We had to download the scrapes in parts, so we need to recombine them.

```{r import and combine gazette}
library("rio")
#Import each dataset
noticesdetail <- rio::import("gazette20k_scrape.csv")
noticespart2 <- rio::import("gazette_next22k_scraped.csv")
noticespart3 <- rio::import("to15apr_detail.csv")
```

We can see that each has a different number of columns, so we need to remove some in order to combine.

```{r show colnames}
colnames(noticespart3)
```

Column 20 is where we get extra info we can remove:

```{r show 20th col}
colnames(noticespart3)[20]
```

What about the second part:

```{r}
head(colnames(noticespart2),30)
```

Let's get rid of that error column:

```{r}
noticespart2 <- noticespart2[,-2]
```

And compare again:

```{r}
head(colnames(noticespart2),30)
head(colnames(noticespart3),30)
```

There's still some mismatch.

Let's just focus on the columns we need.

Check which ones match

```{r}
colnames(noticesdetail) == colnames(noticespart2)
colnames(noticespart2) == colnames(noticespart3)
colnames(noticesdetail) == colnames(noticespart3)
```

The first 8 match, and those are fine - let's just keep those:

```{r reduce to first 19 cols}
noticespart3 <- noticespart3[,1:8]
noticesdetail <- noticesdetail[,1:8]
noticespart2 <- noticespart2[,1:8]
```




And combine

```{r}
#Add another
noticesdetail <- rbind(noticesdetail,noticespart2)

```

And export:

```{r}
write.csv(noticesdetail,"noticesdetail.csv")
```


