---
title: "Insolvencies - Companies House"
output: html_notebook
---

# Insolvencies - Companies House

In two previous notebooks we explored scraped data on insolvency notices. Now we turn our attention to Companies House as a second source of data on the same issue. 

At [this page](http://download.companieshouse.gov.uk/en_output.html) you can download a snapshot of all live companies. 

This is the snapshot for 01/04/2020, split into 6 files.

It's big.

```{r import and combine CH data}
#Import one dataset
chsnapshot010420 <- rio::import("BasicCompanyData-2020-04-01-part1_6.csv")
#Create a list of CSV files that match the pattern
csvfiles <- list.files(pattern="BasicCompanyData-2020-04-01-part[2-6]_6.csv")
#Loop through to create a combined file
for (i in csvfiles){
  onefile <- rio::import(i)
  chsnapshot010420 <- rbind(chsnapshot010420,onefile)
}
```

This is the snapshot for 01/06/2020, split into 6 files.

```{r import and combine CH data June}
#Import one dataset
chsnapshot010620 <- rio::import("BasicCompanyData-2020-06-01-part1_6.csv")
#Create a list of CSV files that match the pattern
csvfiles <- list.files(pattern="BasicCompanyData-2020-06-01-part[2-6]_6.csv")
#Loop through to create a combined file
for (i in csvfiles){
  onefile <- rio::import(i)
  chsnapshot010620 <- rbind(chsnapshot010620,onefile)
}

#Reduce the file 
chsnapshot010620 <- chsnapshot010620[c(1:2,27:30)]
write.csv(chsnapshot010620,"chsnapshot010620.csv")
```

Let's just store 100 to test some techniques with.

```{r store first 100}
chdata100 <- head(onefile, 100)
chdata100
```

There are 4 columns containing SIC codes (because some companies have as many as 4 SIC codes). Let's check the 4th:

```{r check sic values}
table(chdata100$SICCode.SicText_4)
```

98 companies don't have a fourth SIC code, but two do.

We want a simplified data frame which just contains company name and number (for matching) as well as each SIC code it's associated with.

```{r check SIC code colnames}
colnames(chdata100)[27:30]
```

```{r create reduced data frame}
chdata100sic <- chdata100[c(1:2,27:30)]
```

## Create a long data frame

We need `tidyr` to reshape the data and `dplyr` to filter it:

```{r install tidyr and dplyr}
library("tidyr")
library("dplyr")
```

Then we reshape it using `gather` and `filter` out rows without any SIC code:

```{r gather sic codes}
chdata100long <- chdata100sic %>% 
  gather(sicfield, siccode, SICCode.SicText_1:SICCode.SicText_4) %>%
  filter(siccode != "")
#Check how many of each type there are
table(chdata100long$sicfield)
```

So we can see only 15 companies have more than one SIC code, 6 have more than two and there's the two that have 4 SIC codes.

Let's have a look at the SIC codes:

```{r head siccodes}
head(chdata100long$siccode)
```

We need to split the code from the description.


```{r split sic}
chdata100long <- chdata100long %>%
  tidyr::separate(col=siccode, into=c("siccode","sicdesc"), sep=" - ")
head(chdata100long)
```

## Merge with industry division classification

We need the first two digits for a broader industry 'division':

```{r substr sic division}
chdata100long$sicdiv <- substr(chdata100long$siccode,1,2)
```

Now we need to import some data to match with those:

```{r import sic division names} 
sicdivs <- rio::import("sic2007_division_group_class.xls", sheet = 2)
#rename
colnames(sicdivs) <- c("sicdiv","divisionname")
write.csv(sicdivs,"sicdivs.csv")
```

And merge it

```{r merge sic divs}
sicmerged <- merge(chdata100long, sicdivs)
```

There's more detailed categories within each division too - we import those here:

```{r division class import}
#This code is by Anna Khoo
sic_division_group_class <- readxl::read_excel("sic2007_division_group_class.xls")

 

sic_division_group_class <- zoo::na.locf(sic_division_group_class)

sic_division_group_class <- distinct(sic_division_group_class)

 

sic_division_group_class <- sic_division_group_class %>%

  select(1:8)

names(sic_division_group_class) <- c("sector_section", "section_desc", "sector_division","division_desc",

                                     "sector_group", "group_desc", "sector_class", "class_desc")

```

The import doesn't match up all the columns quite right, but we only need them in pairs (where the match is correct) so we can work with those.

First we need to remove periods so they match:

```{r remove periods}
#Replace . with nothing
sic_division_group_class$sector_group<- gsub("[.]","",sic_division_group_class$sector_group)
sic_division_group_class$sector_class<- gsub("[.]","",sic_division_group_class$sector_class)
```

Then split out:

```{r create sectorgroup and class dfs}
colnames(sic_division_group_class)
sicgroups <- sic_division_group_class[5:6]
sicclasses <- sic_division_group_class[7:8]
```

Maybe we can create a cleaner combined dataset (which will also speed up merging) from these:


```{r create combined }
#Copy classes dataset to start with
sicdivclassgroup <- sicclasses
#Extract first 3 digits (group) and 2 digits (division)
sicdivclassgroup$sector_group <- substr(sicclasses$sector_class,1,3)
sicdivclassgroup$sicdiv <- substr(sicclasses$sector_class,1,2)
#Merge descriptions of divs and groups
sicdivclassgroup <- merge(sicdivclassgroup, sicdivs)
sicdivclassgroup <- merge(sicdivclassgroup, sicgroups)
#Remove cols as we only need to match on one
sicdivclassgroup$sicdiv <- NULL
sicdivclassgroup$sector_group <- NULL
```




Let's extract the same digits in the CH data so we can merge:

```{r extract first 3 and 4 digits of SIC codes}
chdata100long$sector_group <- substr(chdata100long$siccode,1,3)
chdata100long$sector_class <- substr(chdata100long$siccode,1,4)
```

Now to merge those - let's repeat the code from earlier where we grabbed the divisions, now we also have the other digits to add, and then merge the other data

```{r merge group and class}
sicmerged <- merge(chdata100long, sicdivs)
sicmerged <- merge(sicmerged, sicgroups)
sicmerged <- merge(sicmerged, sicclasses)
```

## Create a function to do all of this

We now have a dataset which has:

* A row for each SIC code-company combination, allowing us to calculate what percentage of companies fall within a particular code (e.g. out of those 100 companies 30% are in sector B)
* Broader categories added in so we can perform that calculation not just by specific SIC code but SIC division, group and class too

But this was a 100-row sample, not the full 4.5m-row snapshot or even the 257,000-row part.

Let's create a function that combines all the steps and can be run on any dataset.

First we install a package to measure time elapsed:

```{r install tictoc}
install.packages("tictoc")
library(tictoc)
```
Then we create a function:

```{r create function}
chsnapshotcleaner <- function(df){
  print("running chsnapshotcleaner")
  #Remove unnecessary columns
  df <- df[c(1:2,27:30)]
  #wide to long - combine 4 SIC code fields using gather()
  #filter out rows with no SIC code
  #split out code from description
  df <- df %>% 
    tidyr::gather(sicfield, siccode, SICCode.SicText_1:SICCode.SicText_4) %>%
    dplyr::filter(siccode != "") %>%
    tidyr::separate(col=siccode, into=c("siccode","sicdesc"), sep=" - ")
  #create new columns containing just the first 2, 3 and 4 characters of code
  df$sicdiv <- substr(df$siccode,1,2)
  df$sector_group <- substr(df$siccode,1,3)
  df$sector_class <- substr(df$siccode,1,4)
  #merge with descriptions
  df <- merge(df, sicdivclassgroup)
  #return the resulting dataframe to whatever called the function
  return(df)
}
```


I tried the code above 3 times with three different approaches in the function:
* With 3 merges against 3 simple datasets it took 71.9 seconds (using a test df of 257,000)
* With 1 merge against a more complex dataset (it had all 3 codes with the same column names as the df) it took 112.4 secs
* With 1 merge against that dataset with 2 of the 3 codes removed so it only matched against one, it took 57.6 secs

```{r run function and store time taken}
#Start tictoc timer
tictoc::tic("start")
#Run function
dftest <- chsnapshotcleaner(onefile)
#Stop timer and show time
tictoc::toc()
```

Trying it on the full dataset freezes R, however - DON'T run this code:

```{r run function on combined data, eval=FALSE, include=FALSE}
#Start tictoc timer
tictoc::tic("start")
#Run function
chsnapshot010420.sics <- chsnapshotcleaner(chsnapshot010420)
#Stop timer and show time
tictoc::toc()
```

This is another approach - reducing file sizes first:

```{r smaller import}
#Import one dataset
chsnapshot010420 <- rio::import("BasicCompanyData-2020-04-01-part1_6.csv")
chsnapshot010420 <- chsnapshot010420[c(1:2,27:30)]
#Create a list of CSV files that match the pattern
csvfiles <- list.files(pattern="BasicCompanyData-2020-04-01-part[2-6]_6.csv")
#Loop through to create a combined file
for (i in csvfiles){
  onefile <- rio::import(i)
  onefile <- onefile[c(1:2,27:30)]
  chsnapshot010420 <- rbind(chsnapshot010420,onefile)
}
```


Recreate the function with the column-removing line commented out.

```{r}
chsnapshotcleaner <- function(df){
  print("running chsnapshotcleaner")
  #Remove unnecessary columns - commented out as it's now done before running this function
  df <- df[c(1:2,27:30)]
  #wide to long - combine 4 SIC code fields using gather()
  #filter out rows with no SIC code
  #split out code from description
  df <- df %>% 
    tidyr::gather(sicfield, siccode, SICCode.SicText_1:SICCode.SicText_4) %>%
    dplyr::filter(siccode != "") %>%
    tidyr::separate(col=siccode, into=c("siccode","sicdesc"), sep=" - ")
  #create new columns containing just the first 2, 3 and 4 characters of code
  df$sicdiv <- substr(df$siccode,1,2)
  df$sector_group <- substr(df$siccode,1,3)
  df$sector_class <- substr(df$siccode,1,4)
  #merge with descriptions
  df <- merge(df, sicdivclassgroup)
  #return the resulting dataframe to whatever called the function
  return(df)
}
```

And try again:

```{r import with sics}
#Import one dataset
chsnapshot010420b <- rio::import("BasicCompanyData-2020-04-01-part1_6.csv")
chsnapshot010420c <- chsnapshotcleaner(chsnapshot010420b)
#Create a list of CSV files that match the pattern
csvfiles <- list.files(pattern="BasicCompanyData-2020-04-01-part[2-6]_6.csv")
#Loop through to create a combined file
for (i in csvfiles){
  onefilec <- rio::import(i)
  onefilec <- chsnapshotcleaner(onefilec)
  chsnapshot010420c <- rbind(chsnapshot010420c,onefilec)
}
```