---
title: "time_analysis"
author: "Anna Khoo"
date: "29 June 2020"
output: html_document
---
This script is to look at the change in insolvency notices over time.
To avoid double-counting due to mulitple industry/sector matches for around 4500 companies, we will use the 'wide' version of our Companies House-gazette notices data, object chg_postcode_la.

Note, double counting will occur in long version of chg_sic around 4500 entries.
we have to use the wide version, chg_postcode_la, to be able to summarise by count.
```{r setup, include=FALSE}
.libPaths("D:/R/Libraries")
pacman::p_load(tidyverse, zoo, ggplot2, janitor, rio, data.table)
chg_postcode_la <- read_csv("chg_postcode_la.csv")
```

```{r}
pivot_mth_yr <- chg_postcode_la %>%
  group_by(year, month) %>% 
  summarise(insolvencies=n()) %>% 
  arrange(month)

pivot_mth_yr$month <- factor(pivot_mth_yr$month, levels= month.name)
pivot_mth_yr$year <- as.character(pivot_mth_yr$year)

grouped_bars <- ggplot(pivot_mth_yr, 
                       aes(x = month, 
                           y = insolvencies, 
                           fill = year)) +
  geom_bar(stat="identity", position = position_dodge2(preserve = "single")) +
  geom_hline(yintercept = 0, size = 1, colour="#333333")+  
  scale_fill_manual(values = c("#1380A1", "#FAAB18","#8B0000"))+
  theme(axis.text.x = element_text(angle = 30, hjust = 1))+
  labs(title = "liquidation notices", x="Month", y="gazette notices count")
grouped_bars
```
```{r}
grouped_lines <- ggplot(pivot_mth_yr, 
                       aes(x = month, 
                           y = insolvencies, 
                           group  = year))+
  geom_line()+
  geom_point()+
  scale_fill_manual(values = c("#1380A1", "#FAAB18","#8B0000"))+
  theme(axis.text.x = element_text(angle = 30, hjust = 1))+
  labs(title = "liquidation notices", x="Month", y="gazette notices count")
grouped_lines

#nb July data is incomplete in gazette, we only have July 1.
```

```{r}
write_csv(pivot_mth_yr, "insolvency_over_time.csv")
```
```{DON'T RUN r this chunk moved into insolvency_master for central use}
library(lubridate)

chg_postcode_la <- chg_postcode_la %>%
  mutate(date=paste(day, month, year))

chg_postcode_la <- chg_postcode_la %>% 
  mutate(date = dmy(date))
#chg_postcode_la$date <- as.Date(with(chg_postcode_la, paste(year, month, day,sep="-")), "%Y-%m-%d")
```

```{r line with day for granularity (so as to show March did not fall off a cliff)}

#create day-month variable and pivot on that and year
chg_postcode_la <- chg_postcode_la %>% 
  mutate(day_month = substr(date, 6,10))


day_month_year <- chg_postcode_la %>%
    filter(!year==2018) %>% 
    group_by(day_month, year) %>% 
    summarise(insolvencies=n())

day_month_year$day_month <- as.factor(day_month_year$day_month)
day_month_year$year <- as.character(day_month_year$year)

time_series_lines <- ggplot(day_month_year, 
                       aes(x = day_month, 
                           y = insolvencies, 
                           group=year,
                          colour=year))+
  geom_line()+
  geom_point()+
  scale_colour_manual(values = c("#1380A1","#8B0000")) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))+
  labs(title = "liquidation notices", x="Date", y="gazette notices count")
time_series_lines 
```

```{r more or less since lockdown?}

#we can have a play around, checking agg number of notices compared between years
after_lockdown <- chg_postcode_la %>% 
  filter(date>"2020-03-31", date<"2020-07-01")
#corresponding period last year, accounting for weekends, is two days 'ahead' because of leap-year if doing from lockdown
compare_2019 <- chg_postcode_la %>%
  filter(year=="2019") %>% 
  filter(date<"2019-07-01", date>"2019-03-31")

```

```{r March looks high, how much of that is the latter end?}
pivot_mth_yr <- chg_postcode_la %>%
  filter(year==2020, month=="February"|month=="March") %>% 
  group_by(year, month, day) %>% 
  summarise(insolvencies=n())

#pivot_mth_yr$month <- factor(pivot_mth_yr$month, levels= month.name)

grouped_bars <- ggplot(pivot_mth_yr, 
                       aes(x = day, 
                           y = insolvencies, 
                           fill=month)) +
  geom_bar(stat="identity", position = position_dodge2(preserve = "single")) +
  geom_hline(yintercept = 0, size = 1, colour="#333333")+ 
  theme(axis.text.x = element_text(angle = 30, hjust = 1))+
  labs(title = "liquidation notices", x="Day", y="gazette notices count")
grouped_bars

#don't run
group_line <- ggplot(pivot_mth_yr, 
                       aes(x = day, 
                           y = insolvencies)) +
  geom_line() +
  geom_hline(yintercept = 0, size = 1, colour="#333333")+ 
  theme(axis.text.x = element_text(angle = 30, hjust = 1))+
  labs(title = "liquidation notices", x="Day", y="gazette notices count")
group_line


```
Using house_gazette object to get a quick idea of size of liquidated businesses. house_gazette is made using the Insolvency_master.Rmd, the initial match between gazette liquidation notices and Companies House, which contains AccountCategory.
(API call would take a while to fix and add in with that col but as a broad idea:)
```{r house_gazette pivot over time by acc.category}

#we just want company size indication, clean col/subset:
house_gazette$Accounts.AccountCategory <- gsub("TOTAL EXEMPTION ", "",house_gazette$Accounts.AccountCategory)

accounts <- c("FULL","GROUP","MEDIUM","NO ACCOUNTS FILED","SMALL", "DORMANT")

pivot_mth_yr <- house_gazette %>%
  #filter(year==2020) %>%
  filter(Accounts.AccountCategory %in% accounts) %>% 
  group_by(year, month, Accounts.AccountCategory) %>% 
  summarise(insolvencies=n()) %>% 
  arrange(month)


pivot_mth_yr$month <- factor(pivot_mth_yr$month, levels= month.name)
pivot_mth_yr$year <- as.character(pivot_mth_yr$year)

grouped_bars <- ggplot(pivot_mth_yr, 
                       aes(x = month, 
                           y = insolvencies, 
                           fill = year)) +
  facet_wrap(~Accounts.AccountCategory, scales="free")+
  geom_bar(stat="identity", position = position_dodge2(preserve = "single")) +
  geom_hline(yintercept = 0, size = 1, colour="#333333")+  
  theme(axis.text.x = element_text(angle = 30, hjust = 1))+
  labs(title = "liquidation notices", x="Month", y="gazette notices count")
grouped_bars
```

